---
title: "Linear Models"
author: "Derek Lamb"
date: "`r Sys.Date()`"
output: github_document
---
```{r load packages, message = FALSE}
# Load packages
library(tidyverse)
library(p8105.datasets)

# Set default figure options
knitr::opts_chunk$set(
  fig.width = 6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load and clean the Airbnb data
```{r import and tidy}
data("nyc_airbnb")

df_airbnb = nyc_airbnb |> 
  mutate(stars = review_scores_location/2) |> 
  select(price, stars, borough = neighbourhood_group, 
         neighbourhood, room_type) |> 
  filter(borough != "Staten Island")
```

let's fit a model!

```{r basic lm}
fit = df_airbnb |> 
  lm(price ~ stars + borough, data = _)
```

Let's look at the `fit`.

```{r ways of looking at fit}
summary(fit)
summary(fit)$coef
#fitted.values(fit)
```

```{r tidy output}
# Gives model info, r2, AIC, BIC
fit |> 
  broom::glance()

# Reg coefficients
fit |> 
  broom::tidy() |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)
```

## Fit another model

```{r fct reorder on cats}
fit = df_airbnb |> 
  mutate(borough = fct_infreq(borough),
         room_type = fct_infreq(room_type)) |> 
  lm(price ~ stars + borough + room_type, data = _) 

fit |> 
  broom::tidy()
```

## Quick look at diagnostics
```{r model diagnostics}
df_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin() + 
  ylim(-100,500)

#
df_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = stars, y = resid)) + 
  geom_point()

# Follow up about QQ plots
df_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(sample = resid)) + 
  stat_qq() +
  stat_qq_line()

  

```

## Hypothesis test for cat predictor

Fit a "null" and "alternative" model.

```{r hypothesis test on nested model}
fit_null = lm(price ~ stars + borough, data = df_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = df_airbnb)

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Borough-level differences

```{r look at interaction}
fit = df_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _)

fit |> 
  broom::tidy()
```

At least for exploratory analyses, you can separate by borough and fit models, and compare the betas across boroughs.

```{r fit across df}
# Create lm function for nested df
lm_airbnb = function(df){
  lm(
    price ~ stars + room_type, data = df 
  )
}

# Apply lm_airbnb function to each borough
df_airbnb |> 
  nest(df = -borough) |> 
  mutate(
    models = map(df, lm_airbnb),
    results = map(models, broom::tidy)
    ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  knitr::kable(digits = 2)
```

Same thing but just a little different (anonymous function).

```{r anonymous function}
df_airbnb |> 
  nest(df = -borough) |> 
  mutate(
    models = map(df, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)
    ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  knitr::kable(digits = 2)
```

## Binary Outcomes

```{r import data homicide}
df_homicide <- read_csv("data/homicide-data.csv", na = "Unknown") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest")
  ) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```

Fit logistic regression

```{r}
fit_logistic <- df_homicide |> 
  glm(resolved ~ victim_age + victim_race + victim_sex,
      data = _,
      family = binomial())
```

Look at results
```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(or = exp(estimate)) |> 
  select(term, estimate, or)
```

